|

### 2019-09-11 analysis/paper/manuscript.Rmd
First pass at adressing comments from Reviewer 1

  **Abstract**: In the age of big data, scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets. The field of stock assessment is no exception as the volume and variety of the data has grown alongside the computational methods used to integrate them. While this growth in data and model complexity has improved the assessments of many stocks, the process of communicating the results to stakeholders and fisheries managers in a meaningful way has become more daunting. The traditional approach of presenting information across a series of static slides often fails to convey the richness of information available and, as such, important patterns and details are easily overlooked. Here we contend that this problem can be mediated through the effective use of new open source tools for building interactive visualizations. These tools allow a broader audience to conduct detailed explorations of the results, leading to a deeper and collective understanding of both the data and models used to inform stock assessments. As a consequence, the peer review process is more open and accessible and the resulting science advice is improved and widely supported.

@@

Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever-expanding array of data inputs and model outputs. First, we focus on data that are commonly presented in stock assessments and use interactive maps to simplify the detailed exploration of data from a long-term tagging program. Second, we focus on the modeling aspect of stock assessment and demonstrate how dynamic and interactive data visualizations can be used to explore, diagnose and communicate results from an integrated assessment model. This structure corresponds to two important  steps in the stock assessment process: 1) knowing the data, and 2) data modeling.

@@

The data sets used in stock assessments are constantly growing due to the continuation of long-term monitoring efforts, the addition of new monitoring programs, or both. As such, stock assessment biologists need to manage large volumes of data from a variety of sources. Time series of reported landings and catch-at-age are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages. For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality. All of the above-mentioned data sets are complex and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each. This challenge can be mediated, to a degree, by the application of interactive visualization tools as that they allow detailed exploration of the data behind a plot. For instance, the ability to zoom in on features or areas of interest, turn off layers and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of data. This is exemplified by two recently developed interactive tools: 1) a tool designed to quickly examine "fishery independent" survey data; and 2) an interactive  mapping tool developed for the exploration of a long-term tagging study.

@@

@@

@@

The NCAM dashboard (Supplement 2) contains a series of tabs, the first of which provides terse point-form background on the model (tab named "Background"). Subsequent tabs provide a series of diagnostic plots for assessing model fits to the catch ("Catch"), research vessel autumn trawl survey ("RV survey"), Sentinel fishery survey ("SN survey"), inshore acoustic survey ("SS survey"), and tagging ("Tagging") data. For instance, the "RV survey" tab includes plots of observed and predicted values of mean numbers per tow captured in the research vessel survey (Figure \@ref(fig:rv)). The dashboard also includes tabs focused on model estimates such as catchability and selectivity ("Catchability"), stock size and vital rates ("Trends"; Figure \@ref(fig:trends)), and stock productivity ("Productivity"). Finally, some results from a retrospective analyses are included under the "Retro" tab and key inputs and outputs are included as tables under the "Tables" tab. The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents, however, the interactive nature of the plots permit easy and efficient access to significantly more detail. Because the dashboard is contained in an html file, it acts as an interactive assessment document that can be shared for peer-review. This allows colleagues and stakeholders to independently scrutinize details of the data and model that are not easily accessed by users other than the analyst. Such access improves the transparency of the stock assessment model which, in turn, leads to richer discussion and scrutiny of the biological and statistical rigor of the model.

@@

We  believe that the dashboard tool is a step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued. While the entire data-pipeline can be revealed to all by sharing code and data, it is unlikely that an individual has the time to fully review and recreate a particular analysis in its entirety when most participants are already overcommitted [@banks2011]. In the case of NCAM, considerable statistical experience and expertise is required simply to run the model. However, the dashboard removes many of these difficulties. The html file contains all model inputs, outputs and diagnostics in an open and easy to use interactive document, and can be distributed prior to meetings and participants can assess the results and diagnostics at their leisure. Specific experience with the tools used to generate the results are not required to constructively and critically review of the results presented in the dashboard.

@@

We must acknowledge there are costs to adopting the use of interactive tools. The first and obvious one being that staff are required to have both the time and training to effectively develop these tools. In our two examples, the developers (first and second authors) were relatively proficient with R and have natural aptitudes for programming, but had their formal training in field-based population ecology. While some learning time was needed (a few weeks), the learning curve was relatively minor because there was no need to learn a new programming language (e.g. JavaScript, CSS, etc.). Moreover, a growing user base means that most programming issues were readily addressed through simple internet searches. We expect the accessibility and quality of these tools to continue to improve as more people in the fisheries community develop and use interactive documents. It is our experience that the upfront investment in time has already paid dividends in terms of efficiencies of delivering products for subsequent stock assessments, and in the quality of the data exploration and understanding. Scientific exploration is rarely a linear process, and the ability to navigate around a dashboard fits the natural way people pursue ideas.

>>>>

  **Abstract**: Scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets. The field of stock assessment is no exception as the volume and variety of the data has grown alongside the computational methods used to integrate them. While this growth in data and model complexity has improved the assessments of many stocks, the process of communicating the results to stakeholders and fisheries managers in a meaningful way has become more daunting. The traditional approach of presenting information across a series of static slides often fails to convey the richness of information available and, as such, important patterns and details are easily overlooked. Here we contend that this problem can be mediated through the effective use of new open source tools for building interactive visualizations. These tools allow a broader audience to conduct detailed explorations of the results, leading to a deeper and collective understanding of both the data and models used to inform stock assessments. As a consequence, the peer review process is more open and accessible and the resulting science advice is improved and widely supported.

@@

Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever-expanding array of data inputs and model outputs. First, we focus on data that are commonly presented in stock assessments and use interactive plots to simplify the detailed exploration of data from long-term monitoring programs program. Second, we focus on the modeling aspect of stock assessment and demonstrate how dynamic and interactive data visualizations can be used to explore, diagnose and communicate results from an integrated assessment model. This structure corresponds to two important  steps in the stock assessment process: 1) knowing the data, and 2) data modeling.

@@

The data sets used in stock assessments are constantly growing due to the continuation of long-term monitoring efforts, the addition of new monitoring programs, or both. As such, stock assessment biologists need to manage large volumes of data from a variety of sources. Time series of reported landings and catch-at-age are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages. For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality. All of the above-mentioned data sets are complex and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each. This challenge can be mediated, to a degree, by the application of interactive visualization tools as that they allow detailed exploration of the data behind a plot. For instance, the ability to zoom in on features or areas of interest, turn off layers and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of data. This is exemplified by two recently developed interactive tools: 1) a tool designed to quickly examine fishery-independent survey data; and 2) an interactive  mapping tool developed for the exploration of a long-term tagging study.

@@

## Survey data

@@

## Tagging data

@@

The NCAM dashboard (Supplement 2) contains a series of pages, the first of which provides terse point-form background on the model (page named "Background"). Subsequent pages provide a series of diagnostic plots for assessing model fits to catch ("Catch"), survey ("RV survey", "SN survey" and "SS survey"; accessed from the "Surveys" drop-down menu), and tagging ("Tagging") data. For instance, the "RV survey" page includes plots of observed and predicted values of mean numbers per tow captured in the research vessel survey (Figure \@ref(fig:rv)). The dashboard also includes pages focused on model estimates such as catchability and selectivity ("Catchability"), stock size and vital rates ("Trends"; Figure \@ref(fig:trends)), and stock productivity ("Productivity"). Finally, some results from a retrospective analyses are included under the "Retro" page, trends from different models are compared under the "Comps" page, details on the projections are accessed from the "Projections" drop-down menu ("Assumptions", "Past projections", "Retro projections", and "Results" pages), and key inputs and outputs are accessed from "Tables" drop-down menu ("Inputs", "Settings", "Outputs" pages). The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents, however, there are two key benefits of this approach over the *modus operandi* of producing static documents and slides: 1) interactive plots nested in a dashboard permit relatively easy and efficient access to the details as it replaces scrolling through tens, if not hundreds, of pages or slides with mouse-clicks across pages holding data-rich illustrations (i.e. both broader patterns and finer details in the data are accessible via zooming and tooltips); and, 2) the automated nature of the dashboard circumvents the monotonous, time-consuming and error-prone task of copying and pasting figures, tables and values into documents and slides. Both benefits expedite the process of exploring a range of model configurations as the automated output facilitate quick views of standard diagnostics and the interactive plots facilitate detailed explorations and model to model comparisons. The dashboard also makes it easy to share the results with colleagues and stakeholders as it is rendered into a self-contained html file. This allows others to independently scrutinize details of the data and model that are typically only accessible to the analyst. Such access improves the transparency of the stock assessment model which, in turn, leads to richer discussion and scrutiny of the biological and statistical rigor of the model. For instance, visualizations depicting the model assumptions, process errors and confidence intervals around the projections raised important questions on the impact of the assumptions on the projections from the model. These questions were raised during the first assessment meeting in which this tool was used and the ensuing discussion helped the assessment biologists, managers and fisheries representatives in the room gain a deeper understanding of the uncertainties and the patterns in the risk tables [e.g Prob(SSB~2020~ &ge; B~lim~)].

@@

The tools presented here are additions to the growing number of interactive stock assessment tools being developed and applied across the globe (e.g. _________). We  believe that such tools represent an important a step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued. While these tools may not reveal the entire data-pipeline, it is unlikely that an individual has the time to fully review and recreate a particular analysis in its entirety when most participants are already overcommitted [@banks2011]. In the case of NCAM, considerable statistical experience and expertise is required simply to run the model. However, the dashboard removes many of these difficulties. The html file contains all model inputs, outputs and diagnostics in an open and easy to use interactive document that can be distributed prior to meetings and, as such, participants can assess the results and diagnostics at their leisure. Specific experience with the tools used to generate the results are not required to constructively and critically review of the results presented in the dashboard.

@@

We must acknowledge there are costs to adopting the use of interactive tools. The first and obvious one being that staff are required to have both the time and training to effectively develop these tools. In our two examples, the developers (first, second and third authors) were relatively proficient with R and have natural aptitudes for programming, but had their formal training in field-based population ecology. While some learning time was needed (a few weeks), the learning curve was relatively minor because there was no need to learn a new programming language (e.g. JavaScript, CSS, etc.). Moreover, a growing user base means that most programming issues were readily addressed through simple internet searches. We expect the accessibility and quality of these tools to continue to improve as more people in the fisheries community develop and use interactive documents. It is our experience that the upfront investment in time has already paid dividends in terms of efficiencies of delivering products for subsequent stock assessments, and in the quality of the data exploration and understanding. Scientific exploration is rarely a linear process, and the ability to navigate around a dashboard fits the natural way people pursue ideas.
